---
title: "p8105_hw2_cac2225"
author: "Courtney Chan"
date: "October 4, 2018"
output: github_document
---
#Problem 1 NYC Transit Data

First step is to save the relevant dataset to the data subdirectory folder, under the working directory. The tidyverse package is loaded. The data is loaded using a relative path to the subdirectory, with the function readr, and given a name, "nyc_subway_data".

```{r loading library tidyverse}
library(tidyverse)

nyc_subway_data = read_csv(file = "./subdirectory_HW2data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
```

Names of the columns are cleaned using janitor::clean_names, converting names to lower snake case.

```{r data cleaning - clean column names}
names(nyc_subway_data)

nyc_subway_data = janitor::clean_names(nyc_subway_data)
names(nyc_subway_data)
```
Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or  recode function may be useful).