---
title: "p8105_hw2_cac2225"
author: "Courtney Chan"
date: "October 4, 2018"
output: github_document
---
#Problem 1 NYC Transit Data

First step is to save the relevant dataset to the data subdirectory folder, under the working directory. The tidyverse package is loaded. The data is loaded using a relative path to the subdirectory, with the function readr, and given a name, "nyc_subway_data".

```{r loading library tidyverse}
library(tidyverse)

nyc_subway_data = read_csv(file = "./subdirectory_HW2data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
```

Names of the columns are cleaned using janitor::clean_names, converting names to lower snake case.

```{r data cleaning - clean column names}
names(nyc_subway_data)

nyc_subway_data = janitor::clean_names(nyc_subway_data)
names(nyc_subway_data)
```

Preliminarily viewing dataset, the first few rows, last few rows, summary statistics which provide a general overview of whether there is missing column names, data etc.

```{r looking at data}
nyc_subway_data

tail(nyc_subway_data, 5)

skimr::skim(nyc_subway_data)
```

It seems that dataset appears to have all necessary column names, so moving onto the next step of selecting certain columns.

```{r selecting certain columns}
select(nyc_subway_data, line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada)
```

To recode the variable entry using the function if else.

```{r convert entry variable}
nyc_subway_data = mutate(nyc_subway_data, entry = ifelse(entry == "yes", "True", "False"))
```
